---
title: "Scoreboard Vignette"
author: "Mark Barrett"
date: "7/20/2020"
output: html_document
---

```{r setup, include=FALSE}

```

This document outlines how the scoreboard works, and how the processes interact with each other.
We will cover what I will term **Data** processes and **UI** processes, although they do blur together at points.
The **Data** processes relate to the data used, and how data is read, created, and manipulated throughout the course of a game.
The **UI** processes relate to the UI elements, and how the UX utilizes these elements throughout the course of a game.

We will divide either section between processes in one of these phases:

- App Startup
- Game Start
- In-game
- Game Completion

# Packages

To start off, the scoreboard makes use of a number of different packages:

```{r}
# Step into the 'verse
library(tidyverse)
library(lubridate)

# Tnteraction with the database
library(DBI)
library(RPostgres)
library(dbplyr)

# Shiny-related
library(shiny)
library(shinyjs)
library(shinyWidgets)

# Visualization-related
library(gt)
library(extrafont)
```

As outlined in the comments, we use three packages to interact with our database housed by AWS. 
We use three Shiny packages for UI and UX functions.
We use the package-conglomerate tidyverse, `gt`, and `extrafont` for working with data and creating visualizations.

# UI

## App Startup

On startup, we are using a `navbarPage` layout, referencing our custom CSS styling (which seems like it **needs** to be in a `/www` folder to work), and calling ShinyJS.^[I know there is [a ShinyJS article](https://deanattali.com/shinyjs/advanced#usage-navbarpage) about the correct positioning of `useShinyjs` in `navbarPage`, and hopefully we will have updated this by the time anyone is reading this, but in case you're wondering why we're placing the function here - because it works...]

```{r ui-start, eval=F}
ui <- navbarPage(title = "Snappa Scoreboard", id = "navbar", selected = "Player Input",
  theme = "app.css",
  useShinyjs(),
  
#   ...
#   ...
```

### Player Input Screen

Here we make a `tabPanel` for the player inputs.  
Player inputs are created using a custom function called `team_input_ui` which takes team and players arguments. 
In a row below, we create the `start_game` button, as well as `validate_start` which is a reactive UI element that ensures player inputs are neither empty nor duplicated.
Lastly there is a radio button to allow players to choose what score to play to, and some little `helpText` to inform players that they need to enter their names before `start_game` is enabled.
The minimum is two players per team, and we allow for up to 4v4, with no expectation that the teams are balanced i.e. 2v3, 3v4 are allowed.
We will go over the start validation more in the Data Game Start section later on.

Firstly, here is the `team_input_ui` function (contained in the `ui_functions.R` file):

```{r ui-functions}
team_input_ui = function(team, player_choices){
  
  players = str_c("#name_", team, 1:4, "-selectized", collapse = ", ")
  player_inputs = str_c("#name_", team, 1:4, collapse = ", ")
  team_colour = if_else(team == "a", "#e26a6a", "#2574a9")
  
  column(4, align = "center",
         
         wellPanel(
           style = paste("opacity: 0.92; background:", team_colour),
           # Header
           h1(paste("Team", toupper(team)), style = "align: center; color: white; font-size: 600%;"),
           # Player 1
           selectizeInput(paste0('name_', team, '1'), 'Player 1', c(`Player Name`='', player_choices),  options = list(create = TRUE), width = "60%"),
           # Player 2
           selectizeInput(paste0('name_', team, '2'), 'Player 2', c(`Player Name`='', player_choices), options = list(create = TRUE), width = "60%"),
           # Add Player 3 button
           actionBttn(paste0("extra_player_", team, "3"), label = "+ Add Player", style = "unite", color = "danger"), 
           
           # CSS: Increase font size, change color to white, add top and bottom margins
           tags$style(type = "text/css", paste(players, "{color: white; margin-top:30px;margin-bottom:30px;}",
                                               player_inputs, "{color: white; margin-top:30px;margin-bottom:30px;}"))
         )
  )
}

```

The function puts a team name header at the top, creates `selectizeInputs` for player input, and an `actionBttn` (from ShinyWidgets) to allow adding additional players.
The team name argument allows us to name our player inputs based on the team name, and also assign a team colour.
The player inputs use the players already in the database's `players` table, with an option to create new players also.

The style tags should probably be removed and added to our style sheet instead...

Anyway, here is the tabPanel which houses the Player Input screen, using the above function to create the same UI for either team.


```{r player-input, eval=F}
# ...
# ...

tabPanel("Player Input", icon = icon("users"),
         # Fluid Row - 3 columns
         fluidRow(
           team_input_ui("a", pull(players_tbl, player_name)),
           
           # Column 2 - empty
           column(4),
           
           # Column 3 - Team B
           team_input_ui("b", pull(players_tbl, player_name))
           ),
         
         
        # Second row - 3 columns
         fluidRow(
           # Column 1 - empty
           column(4),
           # Column 2
           #    - Start game button
           #    - Score to play to
           column(4,  align = "center",
                  disabled(actionBttn("start_game", 
                                      label = "Throw some dice?", style = "pill", color = "primary")),
                  uiOutput("validate_start"),
                  br(),
                  
                  awesomeRadio(inputId = "play_to", 
                               label = "What score are you playing to?", 
                               choices = list("21" = 1, "32" = 2), 
                               selected = 1, inline = T),
                  br(),
                  helpText("Note: All players must enter their name before the game can begin")
                  ),
           column(4)
           )
),

# ...
# ...

```


## Game Start

Before the game can start, we validate the player inputs.  This is covered in the Data section.
When the game start is validated, users can click the `game_start` button. 
Once they do this, we use the cool functions `insertTab` and `hideTab` to bring the users to the Scoreboard screen.

But first, the UI for a lot of the scoreboard is contained in a function called `team_scoreboard_ui`. 
This function takes two arguments - `left_team` and `right_team` - to enable us to switch the teams to either side of the screen.
The UI consists of a column for Team A, a column for displaying the round and round-tracking buttons, and a column for Team B.

```{r team-scoreboard-ui}
team_scoreboard_ui = function(left_team = "a", right_team = "b"){
  
  team_colours = list("a" = "#e26a6a", "b" = "#2574a9")
  
  div(id = "ScoreboardUI", 
           
           fluidRow(
             # Left Team
             column(width = 4, align = "center",
                     
                     wellPanel(
                       style = paste("opacity: 0.92; background:", team_colours[[left_team]]),
                       # uiOutput("active_die_a"),
                       # Header
                       h1(paste("Team", toupper(left_team)), style = "align: center; color: white; font-size: 550%;"),
                       # Score
                       h2(textOutput(paste0("score_", left_team))),
                       # Score button
                       actionBttn(paste0(left_team, "_score_button"), 
                                  label = "We scored!", color = "danger",
                                  size = "lg"),
                       br(),
                       actionBttn(
                         inputId = paste0("undo_score_", left_team),
                         label = "Undo", style = "unite", color = "danger", icon = icon("undo"), size = "md"
                       ),
                       h3(textOutput(paste0("player_names_", left_team)))
                     )
              ), 
              # Round
              column(width = 4, align = "center",
                     # materialSwitch(
                     #   inputId = "switch_sides",label = "Switch sides", icon = icon("refresh"), 
                     # ),
                     
                     h1("Round", style = "font-size: 600%;"),
                     h3(textOutput("round_num")),
                     fluidRow(actionBttn("previous_round", 
                                         label = "Previous Round", style = "jelly", icon = icon("arrow-left"), color = "primary", size = "lg"),
                              actionBttn("next_round", 
                                         label = "Pass the dice", style = "jelly", icon = icon("arrow-right"), color = "primary", size = "lg")),
                     br(),
              ),
              # Team B
             column(width = 4, align = "center",
                    
                    wellPanel(
                      style = paste("opacity: 0.92; background:", team_colours[[right_team]]),
                      # uiOutput("active_die_a"),
                      # Header
                      h1(paste("Team", toupper(right_team)), style = "align: center; color: white; font-size: 550%;"),
                      # Score
                      h2(textOutput(paste0("score_", right_team))),
                      # Score button
                      actionBttn(paste0(right_team, "_score_button"), 
                                 label = "We scored!", color = "danger",
                                 size = "lg"),
                      br(),
                      actionBttn(
                        inputId = paste0("undo_score_", right_team),
                        label = "Undo", style = "unite", color = "danger", icon = icon("undo"), size = "md"
                      ),
                      h3(textOutput(paste0("player_names_", right_team)))
                    )
             )
              )
  )
}


```

Here is how the whole Scoreboard UI is laid out:

- The overarching `tabPanel` contains a `div`
- Within this div we place a `fluidRow`, the Scoreboard UI created with a custom function, and another div labelled `bottom_buttons`.
- The first `fluidRow` contains an `actionBttn` called `switch_sides` which is used to (you guessed it) switch the team's UI to the other side of the screen.
- A custom function `team_scoreboard_ui` creates the majority of UI elements (above)
- Beneath the scoreboard are buttons for checking `recent_scores`, starting a new game with `new_game`, and finishing the game with `finish_game`.

With all that contained in a function, our remaining Scoreboard UI is created with the following:

```{r game-start-ui, eval=F}
# When we click "Start Game", 
#   - Add new players to the players table
#   - switch to the scoreboard
#   - Set the score outputs and shot number to 0
#   - Record the score we're playing to
#   - Initialize the current game's player_stats table
observeEvent(input$start_game, {

# Add Scoreboard ----------------------------------------------------------

  scoreboard_tab = tabPanel("Scoreboard", icon = icon("window-maximize"), 
                            div(
                              fluidRow(id = "scoreboardrow", 
                                       column(4, align = "left"), 
                                       column(4, align = "center", 
                                              actionBttn("switch_sides", 
                                                         "Switch Sides", style = "unite", color = "primary", icon = icon("refresh"), size = "sm")),
                                       column(4)),
                              team_scoreboard_ui(), 
                              div(id = "bottom_buttons",
                                  fluidRow(
                                    column(width = 4, offset = 4, align = "center",
                                           # Recent Scores
                                           dropdown(
                                             inputId = "recent_scores",
                                             gt_output("recent_scores"),
                                             style = "unite",
                                             size = "lg", 
                                             up = T,
                                             label = "Recent Scores",
                                             icon = icon("backward"),
                                             animate = animateOptions(
                                               enter = animations$fading_entrances$fadeInUp,
                                               exit = animations$fading_exits$fadeOutDown
                                             )
                                           ))
                                    
                                  ),
                                  fluidRow(
                                    column(width = 4, offset = 4, align = "center",
                                           actionBttn("new_game", "Restart game", style = "unite", color = "warning"),
                                           actionBttn("finish_game", "Finish game", style = "unite", color = "warning")
                                    )
                                  )
                                  )
                              
                              )
                            )
  insertTab("navbar", tab = scoreboard_tab, target = "Player Input", position = "after", select = T)  
  hideTab("navbar", "Player Input")
  
#   ...
#   ...

```

## In-game

One of the main drivers of this scoreboard's inception was the lack of an actual scoreboard to keep track of our games' scores.  
Although the data collection aspect is particularly enticing too, we wanted to be able to display the score too.^[And yes, we did have a little flip-book style scoreboard for a while.]

The way we went about displaying this was with reactive outputs. We needed one for both team's scores, and the round number. 
They are pretty simple, just a `renderText` function linked to a value which is reactive.

```{r}
# Output the round number
output$round_num = renderText({
  round_num()
})

# Output Team A's score
output$score_a = renderText({
  vals$current_scores$team_a
})
```


### Recent Scores

Despite having the number of points in front of us, it turned out keeping track of score still wasn't as simple for our working memories as we had hoped.
To solve this, we decided to create a little dropdown which displays recent scores.  
We decided to use the `gt` package and output a single column as a sentence describing the recent score
Admittedly, this could likely be done with `renderText` too, but I was excited to use `gt` and this seemed like a fine use for it!

First up is the `recent_score_sentence` function which converts our data into a sentence:

```{r}
recent_score_sentence = function(scores_data){
  scores_data %>% 
    group_by(score_id) %>% 
    transmute(score_sentence = str_c(player_name, 
                                     " scored ",
                                     points_scored,
                                     " point(s)", 
                                     na.omit(if_else(clink, " with a clink", NA_character_)),  
                                     " for Team ", toupper(scoring_team),
                                     " in round ", round_num, ".",
                                     na.omit(if_else(paddle, str_c(" And it was a", na.omit(if_else(foot, " foot", NA_character_)), " paddle!"), NA_character_)))) %>% 
    ungroup() %>% 
    select(-score_id)
}
```

With this, we pass the top 5 scores (ordered by `score_id`), join these to the `snappaneers()` object to get player names and teams, pass them through the score sentence function, and then to `gt()`.

```{r}
# Recent Scores
output$recent_scores = render_gt({
  
  top_n(vals$scores_db, 5, score_id) %>% 
    arrange(-score_id) %>% 
    left_join(select(snappaneers(), player_id, player_name)) %>% 
    recent_score_sentence() %>% 
    gt() %>% 
    tab_options(column_labels.hidden = T)
})
```



# Data

When we think about the data being used in the Scoreboard, it's important to recognize the different contexts in which the data may exist:

- Database - the data is on the AWS server
- Memory - the data is stored in the app's R environment, an example of which is shown below.
- Reactive Values - the data is being held in the app, but is lost if the app crashes
- Temporary object - the data is being held in the context that it is generated.  This is usually within a function.

## App Startup

```{r connect-to-db}
source("dbconnect.R")

# Round numbers 
rounds = str_c(rep(1:100, each = 2), rep(c("A", "B"), 100))

# Pull db tables for tibble templates
players_tbl = tbl(con, "players") %>% collect()
scores_tbl = tbl(con, "scores") %>% collect()
player_stats_tbl = tbl(con, "player_stats") %>% collect()
game_stats_tbl = tbl(con, "game_stats") %>% collect()
```

On startup, the app connects to the database and initializes the tables in memory.  
We create a `rounds` vector which is used to display the round number of a given shot e.g. the first five rounds are `r rounds[1:5]`.

Meanwhile on the server, we initialize a `reactiveValues` object which contains objects which are used throughout the course of the game. 
These objects are:

- ID values for incrementation
  - `game_id`: Initialized as null, assigned a value later.
  - `new_player_id`: The number of players in the **players** table + 1.
  - `score_id`: Initialized as 0, incremented upon scoring event.
  - `shot_num`: Initialized as 1, incremented on passing the round.
- Database tables (copies)
  - `game_stats`: Game-level data structure. The first 5 columns are pulled on startup, the remainder are generated at game end.
  - `player_stats`: Player-level data structure for the current game. 
  - `players`: Player-level data containing `player_id`. 
  - `scores`: Score-level data structure for recording score events.
- `current_scores`
  - A 2-column tibble containing both teams' scores.
- Rebuttal objects
  - `rebuttal`: a boolean value which identifies if a team is in rebuttal.
  - `rebuttal_tag`: a boolean value which is tagged to identify where teams were in rebuttal in the previous round.
- Objects which identify expected player inputs
  - Because the scoreboard has the option for extra player inputs, we need a mechanism to identify when those player inputs are expected. 
  This came up to prevent empty player inputs from being recorded/causing problems.
- Misc.
  - `score_to`: a numeric value which identifies the score being played to in the current game.
  - `trolls`: records players which have not scored any points (and thus are not in the `scores` table).
  - `switch_counter`: This is a counter which aligns the scoreboard UI when teams switch sides.

The code for creating this object looks like:

```{r}
# In shiny, this object is created with reactiveValues(), here we use a list
#reactiveValues(
vals = list(
    # Initialize new game, player, and score IDs, as well as the shot number
    game_id = NULL,
    new_player_id = sum(dbGetQuery(con, "SELECT MAX(player_id) FROM players"),1),
    score_id = as.integer(0),
    shot_num = as.integer(1),
    
    # DB Tables
    game_stats_db = game_stats_tbl %>% slice(0) %>% select(1:5),
    player_stats_db = player_stats_tbl %>% slice(0),
    players_db = players_tbl,
    scores_db = scores_tbl %>% slice(0),

    # dataframe of the players and their teams
    # Current Scores
    current_scores = tibble(
      team_a = 0,
      team_b = 0
    ),
    
    rebuttal = NULL,
    rebuttal_tag = F, 
    
    

    # Values used in scoring events
    score = NULL,
    error_msg = NULL,
    print = FALSE,
    
    score_to = NULL,
    
    # Holds the trolls (more for simplicity of code
    # than direct need)
    trolls = NULL,
    
    #Records when the extra player ui's are open and 
    # allows the app to pay attention to empty strings
    # only during select times
    want_a3 = F,
    want_a4 = F,
    want_b3 = F,
    want_b4 = F,
    
    switch_counter = 1
  )
```

We also initialize a few little utility values:

- A round counter which indexes our earlier `rounds` vector
- A list of active player inputs
- A tibble containing the player names, IDs, and respective teams of the current players
- A reactive counter of active players

```{r reactives, eval=F}

# Increment round number
round_num = reactive({
  rounds[vals$shot_num]
  })
  
# Active input buttons
#   - List of player inputs which are not null
active_player_inputs = reactive({
  list("a1" = input$name_a1, "a2" = input$name_a2, "a3" = input$name_a3, "a4" = input$name_a4, 
       "b1" = input$name_b1, "b2" = input$name_b2, "b3" = input$name_b3, "b4" = input$name_b4) %>% 
    discard(is_null)
})

# Snappaneers - | Team | Player name | Player ID  |
snappaneers = reactive({
  
  tibble(
    # Team pulls the first letter from their input name
    team = str_extract(names(active_player_inputs()), ".{1}"),
    player_name = active_player_inputs() %>% flatten_chr()
  ) %>% 
    # Remove empty player inputs
    filter(player_name != "") %>% 
    left_join(vals$players_db, by = "player_name") %>% 
    # Add shot count
    add_shot_count()
})


# Length of active player inputs
num_players = reactive({
  length(active_player_inputs()[active_player_inputs() != ""])
})
```


## Game Start

There are two processes which occur when a user clicks **Throw some dice** - a validation of player inputs, and the initialization of objects used throughout the game.

### Validation of player inputs

There is only one condition required for a game to start - **No input can be an empty string**.

Here at SnappaMetrics, we like to consider ourselves pretty open-minded but unfortunately we are not open to names which have no characters.
No doubt, Elon Musk will invalidate our intolerance with a new child called `""` in a few years, but until then we are using the `shinyjs::disable` function to disable the `game_start` button when there are empty player inputs.
We use `shiny::validate` to provide a useful error message while any player inputs remain empty.

Notice that this object is an **output**.
It is an output because we want to use the side effects of the `validate` function to provide the error message.
The visual output of this object come from the `need` calls inside the `validate` function.


```{r game-start-validate, eval=F}

# Create a UI output which validates that there are four players and the names are unique
output$validate_start = reactive({
  # If one of the first two players on each team is removed, disable the button again.
  # This goes above the validate check because  it needs to be updating before the validate
  # check is failed, or else the logic isn't going to pass through
  
  if(any(input$name_a1 == "",
         input$name_a2 == "",
         input$name_b1 == "",
         input$name_b2 == "")){
    shinyjs::disable("start_game")
  }
  
  validate(
    need(input$name_a1 != "", label = "Player A1"),
    need(input$name_a2 != "", label = "Player A2"), 
    need(input$name_b1 != "", label = "Player B1"), 
    need(input$name_b2 != "", label = "Player B2")
    )
  
  #Record the players that you need to be looking for
  # (i.e., which ui elements are open right now?)
  
  
  # If the number of unique snappaneer names is the same as the number of active player inputs
  #   => enable start button
  
  if(sum(length(unique(snappaneers()$player_name)), 
         sum(
           c(isTRUE(active_player_inputs()$a3 == "" & vals$want_a3), 
           isTRUE(active_player_inputs()$a4 == "" & vals$want_a4), 
           isTRUE(active_player_inputs()$b3 == "" & vals$want_b3), 
           isTRUE(active_player_inputs()$b4 == "" & vals$want_b4)) 
           )
         ) == num_players()){ 
    shinyjs::enable("start_game")
  } 
  
  # If the number of unique snappaneer names is not the same as the number of active player inputs
  #   => disable start button
  if(sum(length(unique(snappaneers()$player_name)), 
         sum(
           c(isTRUE(active_player_inputs()$a3 == "" & vals$want_a3), 
              isTRUE(active_player_inputs()$a4 == "" & vals$want_a4), 
              isTRUE(active_player_inputs()$b3 == "" & vals$want_b3), 
              isTRUE(active_player_inputs()$b4 == "" & vals$want_b4)
              ) 
         )
  ) != num_players()){ 
    
  shinyjs::disable("start_game")
  }
  

  })
```

### `start_game`

After the game start validation, the `start_game` button is clickable. 
`observeEvent` detects that the button has been clicked, `insertTab` adds the Scoreboard UI, and the player input page is no longer visible.
If there are any players which are not already in our database, we use `purrr::iwalk` to add them (and subsequently increment `new_player_id`).

Then each team's values in `current_scores` are set to 0. (TODO: This can be done in `restart_game` instead)
The `game_id` and `score_to` (the score threshold) are set, and the `game_stats`, `player_stats`, and `scores` tables are set to a skeleton version of each table (using `slice(df, 0)`).

```{r game-start, eval=F}
# When we click "Start Game", 
#   - Add new players to the players table
#   - switch to the scoreboard
#   - Set the score outputs and shot number to 0
#   - Record the score we're playing to
#   - Initialize the current game's player_stats table
observeEvent(input$start_game, {

# Add Scoreboard ----------------------------------------------------------

  scoreboard_tab = tabPanel("Scoreboard", icon = icon("window-maximize"), 
                            div(
                              fluidRow(id = "scoreboardrow", 
                                       column(4, align = "left"), 
                                       column(4, align = "center", 
                                              actionBttn("switch_sides", 
                                                         "Switch Sides", style = "unite", color = "primary", icon = icon("refresh"), size = "sm")),
                                       column(4)),
                              team_scoreboard_ui(), 
                              div(id = "bottom_buttons",
                                  fluidRow(
                                    column(width =4, offset = 4, align = "center",
                                           # Recent Scores
                                           dropdown(
                                             inputId = "recent_scores",
                                             gt_output("recent_scores"),
                                             style = "unite",
                                             size = "lg", 
                                             up = T,
                                             label = "Recent Scores",
                                             icon = icon("backward"),
                                             animate = animateOptions(
                                               enter = animations$fading_entrances$fadeInUp,
                                               exit = animations$fading_exits$fadeOutDown
                                             )
                                           ))
                                    
                                  ),
                                  fluidRow(
                                    column(width = 4, offset = 4, align = "center",
                                           actionBttn("new_game", "Restart game", style = "unite", color = "warning"),
                                           actionBttn("finish_game", "Finish game", style = "unite", color = "warning")
                                    )
                                  )
                                  )
                              
                              )
                            )
  insertTab("navbar", tab = scoreboard_tab, target = "Player Input", position = "after", select = T)  
  hideTab("navbar", "Player Input")
  

  # Add new players to the players table
  iwalk(snappaneers()$player_name, function(die_thrower, index){
    # If the player is not in the players table
    if(!(die_thrower %in% vals$players_db$player_name)){
      
      # Add a row to the players table with the new player's name and new ID
      vals$players_db = bind_rows(vals$players_db,
                               tibble(
                                 player_id = vals$new_player_id,
                                 player_name = die_thrower))
      
      # Increment the ID for the next new player
      vals$new_player_id = vals$new_player_id+1
      
    } else {
      invisible()
    }
  })
  
  # Set the score outputs and shot number to 0
  vals$current_scores$team_a = 0
  vals$current_scores$team_b = 0
  vals$scores_db = slice(scores_tbl, 0)
  vals$game_id = as.integer(sum(dbGetQuery(con, "SELECT MAX(game_id) FROM game_stats"),1 , na.rm = T))
  
  vals$game_stats_db = bind_rows(vals$game_stats_db,
            tibble(
              game_id = vals$game_id,
              num_players = nrow(snappaneers()),
              game_start = as.character(now(tzone = "America/Los_Angeles")),
              game_end = NA_character_,
              night_dice = NA
            ))
  
  # Record the score we're playing to
  vals$score_to = case_when(input$play_to == 1 ~ 21,
                            input$play_to == 2 ~ 32)
  
  
  # Initialize the current game's player_stats table
  vals$player_stats_db = slice(vals$player_stats_db, 0)

})

```


## In-game

Once we are in-game, users have the option to:

- Keep track of the round with `next_round` / `previous_round`
- Score for either team using `[TEAM]_score_button`
- Undo their team's most recent score using `undo_score_[TEAM]`
- Restart the game using `new_game`
- Finish the game using `finish_game`

### Round tracking

Passing the round or returning to the previous round allows users to accurately track the round they are in. 
The process is relatively simple, using `observeEvent` the round is incremented upon pressing `next_round` and decremented upon pressing `previous_round`.
You will notice the event associated with `next_round` has a lot more going on.  
The round number is used to check whether either team is in rebuttal, and to bring up a dialogue box when the game has reached its scoring threshold.

```{r round-tracking, eval=F}
# When next round button is pushed
observeEvent(input$next_round, {
  if (vals$rebuttal_tag == T){
    if (vals$rebuttal == T){
      click("finish_game")
    } else {
      vals$rebuttal_tag = F
    }
  } else{
  }
  
  vals$shot_num = vals$shot_num+1

  vals$rebuttal = rebuttal_check(a = vals$current_scores$team_a, b = vals$current_scores$team_b,
                                 round = round_num(), points_to_win = vals$score_to)
    
  if (vals$rebuttal == T) {
    vals$rebuttal_tag = T
    showNotification(str_c("Rebuttal: ", "Team ", 
                           str_sub(round_num(), start = -1),
                           " needs ", str_c(
                             abs(vals$current_scores$team_a - vals$current_scores$team_b) - 1),
                           " points to bring it back"
                           )
                    ) 
  } else {
  }
    
  })


# When previous round button is pushed
observeEvent(input$previous_round, {
  validate(
    need(vals$shot_num > 1, label = "Can't go below 0", message = "It's the first round still")
  )
  vals$shot_num = vals$shot_num-1
})

```


### Scoring

Programming the scoring was somewhat complex due to the nature of our rules.
We are proud evangelists of the phrase *the rules protect the fun*, and as such we have engineered a lot of fun protection into this game.

First, users must click the **score_button** for their team, located beneath the score for either team.
Reacting to this event, we determine a list of `eligible_shooters` to determine which names to give to the `score_check` function to present as options when selecting the player that scored.

```{r scoring1, eval=F}
observeEvent(input$a_score_button, {
  vals$error_msg <- NULL
  
  eligible_shooters = filter(snappaneers(), team == "a") %>% 
    pull(player_name) %>% 
    sample()
  
  showModal(
    score_check(team = "a", 
                players = eligible_shooters))
})

```

We will go over `score_check` more in-depth in the UI section because it generates the UI used to enter scores, but it is important to know 

```{r scoring2, eval=F}

# Team A presses score button
observeEvent(input$ok_a, {
  # validate(
  #   need(input$score < 8, label = "C'mon, you did not score that many points")
  # )
  # Check that the round/shooter combination makes sense / indicated a paddle
  validate(
    need(
      any(
        # Typical Offense
        str_detect(rounds[vals$shot_num], "[Aa]") & 
          str_detect(pull(filter(snappaneers(), player_name == input$scorer), team), "[Aa]"),
        # Typical Paddle
        str_detect(rounds[vals$shot_num], "[Bb]") &
          str_detect(pull(filter(snappaneers(), player_name == input$scorer), team), "[Aa]") & 
          input$paddle == T,
        # Somebody messed up on the other team
        str_detect(rounds[vals$shot_num], "[Aa]") &
        str_detect(pull(filter(snappaneers(), player_name == input$scorer), team), "[Bb]") & 
          input$paddle == T),
      message = "That entry doesn't make sense for this round/shooter combination"),
    if (pull(filter(snappaneers(), player_name == input$scorer), player_id) %in%
        pull(filter(vals$scores_db, round_num == round_num() & paddle == F), player_id)){
      need(input$paddle == T,
           message = "That person has already scored a non paddle point this round")
      
    }
  )  

  # set score
  score = as.integer(input$score)
  vals$score <- score
  
  
  # Check score i not null, remove the dialog box
  if (!is.null(vals$score)) {
    removeModal()
    vals$print <- TRUE
    
    # Update the team score
    vals$current_scores$team_a = vals$current_scores$team_a + vals$score
    
    # Increment the score_id
    vals$score_id = as.integer(vals$score_id+1)
    
    ## Identify scoring characteristics
    # Player ID
    scorer_pid = pull(filter(vals$players_db, player_name == input$scorer), player_id)
    # Were they shooting?
    scorers_team = pull(filter(snappaneers(), player_name == input$scorer), team) # pull the scorer's team from snappaneers
    shooting_team_lgl = all(str_detect(round_num(), "[Aa]"), scorers_team == "a") # Are they on team A & did they score for team A?
    
    # Add the score to the scores table
    vals$scores_db = bind_rows(vals$scores_db,
                               tibble(
                                 score_id = vals$score_id,
                                 game_id = vals$game_id,
                                 player_id = scorer_pid,
                                 scoring_team = "a",
                                 round_num = round_num(),
                                 points_scored = score,
                                 shooting = shooting_team_lgl,
                                 paddle = any(input$foot, input$paddle),
                                 clink = input$clink,
                                 foot = input$foot
                               ))
    

    # Update game stats table
    vals$player_stats_db = vals$scores_db %>% 
      # Join scores to snappaneers to get each player's team
      left_join(snappaneers(), by = "player_id") %>% 
      # Group by game and player, (team and shots are held consistent)
      group_by(game_id, player_id, team, shots) %>% 
      # Calculate summary stats
      summarise(total_points = sum(points_scored),
                ones = sum((points_scored == 1)),
                twos = sum((points_scored == 2)),
                threes = sum((points_scored == 3)),
                impossibles = sum((points_scored > 3)),
                paddle_points = sum(points_scored* (paddle | foot)),
                clink_points = sum(points_scored*clink),
                points_per_round = total_points / last(shots),
                off_ppr = sum(points_scored * !(paddle | foot))/ last(shots),
                def_ppr = paddle_points/last(shots),
                toss_efficiency = sum(!(paddle | foot ))/last(shots)) %>% 
      ungroup()
    
    # Congratulate paddlers
    if(input$paddle & str_detect(pull(filter(snappaneers(), player_name == input$scorer), team), "[Aa]") ){
      showNotification("That's some hot shit!", id = "paddle")
    }
    if(input$paddle & str_detect(pull(filter(snappaneers(), player_name == input$scorer), team), "[Bb]") ){
      showNotification("It's a bold strategy Cotton, let's see if it pays off for them.")
    }
  } else {
    vals$error_msg <- "You did not input anything."
  }
  
  # If the game is in rebuttal, remind players
  # of the points needed to bring it back
  vals$rebuttal = rebuttal_check(vals$current_scores$team_a, 
                                 vals$current_scores$team_b,
                                 round_num(), vals$score_to)
  
  #    if (!is.null(vals$rebuttal)) {
  if (vals$rebuttal == T & vals$rebuttal_tag == T) {
    showNotification(str_c("Rebuttal: ", "Team ", 
                           str_sub(round_num(), start = -1),
                           " needs ", str_c(
                             abs(vals$current_scores$team_a - vals$current_scores$team_b) - 1),
                           " points to bring it back"
    )
    )
  } else {
    
  }
  validate(
    need((vals$current_scores$team_a == 18 && vals$current_scores$team_b == 12) || (vals$current_scores$team_a == 12 && vals$current_scores$team_b == 18), label = "eighteen_twelve")
  )
  
  
  sendSweetAlert(session, 
                 title = "1812",
                 text = "Everyone roll a die, lowest roll takes a shot.",
                 type = "warning")

})

```



